Do I Need to Model You in order to Learn With You?

### 2015-04-22 mutual-modeling.tex
Split Study 5 picture in two
Several minor fix-up

    \item $\mathcal{H}_{2}$: $\mathcal{M}^{\circ}(A,B)$ depends upon  $B$'s ability of
        effort to help $A$ to model him/herself,

@@

This leads to two \emph{triangle questions} (Fig.~\ref{mm_triangles}): Do $A$ and
B have the same accuracy when modelling C? $\Delta_2 = \Delta(\M{A}{C}{X},
\M{B}{C}{X})$. If it is the same, $\mathcal{H}_{2}$ and
$\mathcal{H}_{3}$\fxfatal{why H3?} gain
over $\mathcal{H}_{1}$

Conversely, does $C$ model more accurately $A$ than B? $\Delta_3= \Delta(\M{C}{A}{X},
\M{C}{B}{X})$ A positive answer would support $\mathcal{H}_{3}$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Study 1: Effect of an awareness tool on \gModel{A}{B} in a virtual
game}

@@

\begin{figure*}
        \centering
        \includegraphics[width=0.9\textwidth]{image4.png}
        \caption{Screenshot of the {\sc SpaceMiners} game}
        \label{study1:spaceminer}
\end{figure*}

@@

\begin{figure}[ht!]
        \centering
        \includegraphics[width=\columnwidth]{image5.png}
        \caption{\gmodel{A}{A} and \gmodel{A}{B} questionnaires in SpaceMiners
        (translated from French).}

@@

\paragraph{Study-specific questions} The awareness tool permitted higher group
performance, but it did not improve the accuracy of the mutual model. Since
teams were free to use the awareness tool or not (the \emph{scout} mode), we
performed a post-hoc split of players depending on how much time they used it.
The split point was the mean of time spent in the scout mode and it led to the
constitution of two groups made up of 12 individuals ``short time in scout
mode'' and 24 individuals ``long time in scout mode''\fxfatal{Inconsistent! 36
participants in awareness condition??}. A two-way analysis of
variance conducted on these contrasted groups revealed that pairs in the
awareness condition who spent more time in the scout mode reached higher levels
of \gModel{A}{B}($F = 8.02, p = 0.015$). Of course, a \textit{post-hoc} split
does not support a causal direction. An alternative explanation could be that
good modellers are more social and hence appreciate the awareness tool.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Study 2: Effect of an awareness tool on \gModel{A}{B}  in a pervasive
game}

@@

\begin{figure*}[h!t]
        \centering
        \begin{subfigure}{.3\textwidth}
            \includegraphics[width=\linewidth]{image6.png}
            \caption{A's drawing of $B$'s path.}
        \end{subfigure}
        \begin{subfigure}{.3\textwidth}
            \includegraphics[width=\linewidth]{image7.png}
            \caption{Actual path followed by $B$.}
        \end{subfigure}
        \caption{Reported and actual path of one of the player, during the {\sc
        CatchBob} game.}
        \label{study2:paths}
\end{figure*}

@@

\paragraph{Triangle questions} Regarding $\Delta_2$, the correlation between
\gModel{A}{B} and \gModel{A}{C} is significant: $r=0.30, p <.01$. Concerning
$\Delta_3$, the correlation between \gModel{A}{C} and \gModel{B}{C} is
significant as well: $r=0.43, p <.001$\fxfatal{Attention! $\Delta_2$ and
$\Delta_3$ seem to be inverted! Cf fig 2}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Study 3:  Effect of media richness on \gModel{A}{B} in argumentation}

@@

\paragraph{Grounding criterion} The maps produced by teams were
ranked\fxfatal{How are they ranked?}. We used
the average rank as a estimation of the team performance and correlated it with the
average of the 6 values of \gModel{A}{B} per team (\gModel{A}{B}, \gModel{A}{C},
\gModel{B}{A},...): The correlation is -0.22 \fxfatal{p value?}: teams with a good \gModel{A}{B}
team to be better ranked.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Study 4:  Effects of a script on \gModel{A}{B}  in concept mapping}

@@

\begin{multline} \label{eq:study4.1}
    \Mdeg{A}{B}{knowledge} = \\
        \sum_{i=1}^{3} (\M{A}{B}{chapter_i} - score(B, chapter_i))
\end{multline}

\begin{multline} \label{eq:study4.2}
    \Mdeg{A}{A}{knowledge} = \\
        \sum_{i=1}^{3}  (\M{A}{A}{chapter_i} - score(A, chapter_i))
\end{multline}

\begin{multline} \label{eq:study4.3}
    \Mdeg{A}{B}{knowledge} = \\
        \sum_{i=1}^{3}  (\M{A}{B}{chapter_i} - \M{B}{A}{chapter_i})
\end{multline}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Study 5: \Model{A}{B}{knowledge} in concept mapping}

@@

\begin{figure*}
        \centering
        \includegraphics[width=0.9\textwidth]{image11.png}
        \caption{Screenshot of the KAT condition during the concept-map building
        phase.}
        \label{study5:kat}
\end{figure*}

@@

\paragraph{Grounding criterion} A linear regression revealed a positive relation
between \gModel{A}{B} and the collaborative learning gain\fxfatal{What is the
collaborative learning gain} of the pair $\{A,
B\}$: $\beta= .401, p < .001, r2adj. =.15$, large effect. This relationship was
not significant in the previous experiment which as been conducted with the same
task: using the same task should produce the same grounding criterion. This
probably due to the fact that \gModel{A}{B} was influenced by the KAT.

@@

\begin{table*}[h!t]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{p{3cm}|p{4cm}|p{4cm}|p{4cm}|p{4cm}|p{4cm}}
\textit{}                     & \textbf{Study 1}                 & \textbf{Study 2}                & \textbf{Study 3}          & \textbf{Study 4}          & \textbf{Study 5}             \\ \hline
\textit{Task}                 & Game in virtual Space            & Game in physical Space          & Building an argument map  & Building a concept map    & Building a concept map       \\
\textit{Interactions}         & Audio                            & Written                         & Audio / Video             & Audio                     & Audio                        \\
\textit{Shared editor}        & 3D Space                         & 2D Map                          & Concept Map               & Concept Map               & Concept Map                  \\
\textit{Group Size}           & 2                                & 3                               & 3                         & 3                         & 3                            \\
\textit{Duration (mean)}      & 90min                            & 16min                           & 61min                     & 90min                     & 90min                        \\
\textit{Awareness tool}       & Partner's position / orientation & Partner's current/past position & -                         & Partner's concept map     & Partner's scores at pre-test \\
\textit{Dependent var.}   & Team/Ind. performance            & Team/Ind. performance           & Team performance          & Team/Ind. knowledge gains & Team/Ind. knowledge gains    \\
    \textit{Independent var.} & Awareness tool \textit{vs} not.           & Awareness tool \textit{vs}  not.          & Audio \textit{vs} Audio+Video & Script \textit{vs} not.            & Awareness tool \textit{vs} not.       \\
\textit{MM°-DV correlation}   & +0.42                            & +0.15                           & +0.22                     & +0.08                     & +0.40                        \\
\textit{$\Delta_1$}                   & +0.38 (significant)                      & +0.41 (significant)                     & ???                       & +0.26 (near significant)          & +0.05 (ns)              \\
\textit{$\Delta_2$}                   &                                  & +0.30                           & +0.43                     &                           &                              \\
\textit{$\Delta_3$}                   &                                  & ???                             & +0.66                     &                           &                              \\
\textit{$\Delta_4$}                   &                                  &                                 &                           & +0.05                     &                             
\end{tabular}
}
\caption{Comparison of the studies}
\label{synthesis_table}
\end{table*}

Coming back to our initial hypotheses, the $\Delta_2$ results in studies 1, 2
and 3\fxfatal{No $\Delta_2$ in study 1! Dyads!} reinforce both $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The
\emph{interaction hypothesis} $\mathcal{H}_{3}$ was that the activity of
modelling one's partner is not simply the result of an individual cognitive
effort but overall the side-effect of the quality of interactions. The
\emph{second-order-modelling hypothesis} $\mathcal{H}_{2}$ stated that, if $A$ is
accurate in modelling $B$, $A$ should notice when $B$ misunderstands $A$ and should be
able to repair. A good modeler would somehow drive the modelling of his or her
partner.  This \emph{drive} will however occur through verbal cues and repairs
and is hence not easy to differentiate from $\mathcal{H}_{3}$.  The
contradictory $\Delta_1$ results in studies 4 (to some extent) and 5 weaken the
hypothesis.

>>>>

    \item $\mathcal{H}_{2}$: $\mathcal{M}^{\circ}(A,B)$ depends upon  $B$'s ability or
        effort to help $A$ to model him/herself,

@@

This leads to two \emph{triangle questions} (Fig.~\ref{mm_triangles}): Do $A$ and
$B$ have the same accuracy when modelling $C$ ($\Delta_2 =
\Delta(\mathcal{M}^{\circ}(A,C,X),
\mathcal{M}^{\circ}(B,C,X))$)? If it is the same, $\mathcal{H}_{2}$ gains over
$\mathcal{H}_{1}$

Conversely, does $C$ model more accurately $A$ than $B$ ($\Delta_3=
\Delta(\mathcal{M}^{\circ}(C,A,X),
\mathcal{M}^{\circ}(C,B,X))$)? A positive answer would support $\mathcal{H}_{3}$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{{\bf Study 1}: Effect of an awareness tool on \gModel{A}{B} in a virtual
game}

@@

\begin{figure*}
        \centering
        \includegraphics[width=\textwidth]{image4.png}
        \caption{Screenshot of the {\sc SpaceMiners} game}
        \label{study1:spaceminer}
\end{figure*}

@@

\begin{figure}[ht!]
        \centering
        \includegraphics[width=0.7\columnwidth]{image5.png}
        \caption{\gmodel{A}{A} and \gmodel{A}{B} questionnaires in SpaceMiners
        (translated from French).}

@@

\paragraph{Study-specific questions} The awareness tool permitted higher group
performance, but it did not improve the accuracy of the mutual model. Since
teams were free to use the awareness tool or not (the \emph{scout} mode), we
performed a post-hoc split of players depending on how much time they used it.
The split point was the mean of time spent in the scout mode and it led to the
constitution of two groups made up of 12 individuals ``short time in scout
mode'' and 24 individuals ``long time in scout mode''. A two-way analysis of
variance conducted on these contrasted groups revealed that pairs in the
awareness condition who spent more time in the scout mode reached higher levels
of \gModel{A}{B}($F = 8.02, p = 0.015$). Of course, a \textit{post-hoc} split
does not support a causal direction. An alternative explanation could be that
good modellers are more social and hence appreciate the awareness tool.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{{\bf Study 2}: Effect of an awareness tool on \gModel{A}{B}  in a pervasive
game}

@@

\begin{figure*}[h!t]
        \centering
        \begin{subfigure}{.4\textwidth}
            \includegraphics[width=\linewidth]{image6.png}
            \caption{A's drawing of $B$'s path.}
        \end{subfigure}
        \begin{subfigure}{.4\textwidth}
            \includegraphics[width=\linewidth]{image7.png}
            \caption{Actual path followed by $B$.}
        \end{subfigure}
        \caption{Reported and actual path of one of the player, during the {\sc
        CatchBob} game.}
        \label{study2:paths}
\end{figure*}

@@

\paragraph{Triangle questions} Regarding $\Delta_2$, the correlation between
\gModel{A}{C} and \gModel{B}{C} is significant: $r=0.43, p <.001$. Concerning
$\Delta_3$, the correlation between \gModel{A}{B} and \gModel{A}{C} is
significant as well: $r=0.30, p <.01$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{{\bf Study 3}:  Effect of media richness on \gModel{A}{B} in argumentation}

@@

\paragraph{Grounding criterion} The maps produced by teams were ranked by three
independent judges on completeness and structure quality (Kendall's $W=0.474$,
limited agreement). We used the average rank as a estimation of the team
performance and correlated it with the average of the 6 values of \gModel{A}{B}
per team (\gModel{A}{B}, \gModel{A}{C}, \gModel{B}{A},...). The correlation is
-0.22: teams with a good \gModel{A}{B} tend to be better ranked.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{{\bf Study 4}:  Effects of a script on \gModel{A}{B}  in concept mapping}

@@

\begin{multline} \label{eq:study4.1}
    \Mdeg{A}{B}{knowledge} = 
        \sum_{i=1}^{3} (\M{A}{B}{chapter_i} - score(B, chapter_i))
\end{multline}

\begin{multline} \label{eq:study4.2}
    \Mdeg{A}{A}{knowledge} = 
        \sum_{i=1}^{3}  (\M{A}{A}{chapter_i} - score(A, chapter_i))
\end{multline}

\begin{multline} \label{eq:study4.3}
    \Mdeg{A}{B}{knowledge} = 
        \sum_{i=1}^{3}  (\M{A}{B}{chapter_i} - \M{B}{A}{chapter_i})
\end{multline}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{{\bf Study 5}: \Model{A}{B}{knowledge} in concept mapping}

@@

\begin{figure*}[h!t]
        \centering
        \begin{subfigure}{.6\textwidth}
            \includegraphics[width=\linewidth]{study5-conceptmap.png}
        \end{subfigure} \\
        \begin{subfigure}{.7\textwidth}
            \includegraphics[width=\linewidth]{study5-kat.png}
        \end{subfigure}
        \caption{Screenshot of the KAT condition during the concept-map building
        phase.}
        \label{study5:kat}
\end{figure*}

@@

\paragraph{Grounding criterion} A linear regression revealed a positive relation
between \gModel{A}{B} and the learning gain of the pair $\{A, B\}$ (calculated
as the average of the individual learning gains): $\beta= .401, p < .001, r2adj.
=.15$, large effect. This relationship was not significant in the previous
experiment which as been conducted with the same task (using the same task
should produce the same grounding criterion). This is probably due to the fact that
\gModel{A}{B} was influenced by the KAT.

@@

\begin{table*}[h!t]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{p{3cm}|p{4cm}|p{4cm}|p{4cm}|p{4cm}|p{4cm}}
\textit{}                     & \textbf{Study 1}                 & \textbf{Study 2}                & \textbf{Study 3}          & \textbf{Study 4}          & \textbf{Study 5}             \\ \hline
\textit{Task}                 & Game in virtual space            & Game in physical space          & Building an argument map  & Building a concept map    & Building a concept map       \\
\textit{Interactions}         & Audio                            & Written                         & Audio / Video             & Audio                     & Audio                        \\
\textit{Shared editor}        & 3D Space                         & 2D Map                          & Concept Map               & Concept Map               & Concept Map                  \\
\textit{Group Size}           & 2                                & 3                               & 3                         & 3                         & 3                            \\
\textit{Duration (mean)}      & 90min                            & 16min                           & 61min                     & 90min                     & 90min                        \\
\textit{Awareness tool}       & Partner's position / orientation & Partner's current/past position & -                         & Partner's concept map     & Partner's scores at pre-test \\
\textit{Dependent var.}   & Team/Ind. performance            & Team/Ind. performance           & Team performance          & Team/Ind. knowledge gains & Team/Ind. knowledge gains    \\
    \textit{Independent var.} & Awareness tool \textit{vs} not.           & Awareness tool \textit{vs}  not.          & Audio \textit{vs} Audio+Video & Script \textit{vs} not.            & Awareness tool \textit{vs} not.       \\
\textit{MM°-DV correlation}   & +0.42                            & +0.15                           & +0.22                     & +0.08                     & +0.40                        \\
\textit{$\Delta_1$}           & +0.38 (significant)              & +0.41 (significant)             & ???                       & +0.26 (near significant)          & +0.05 (ns)              \\
\textit{$\Delta_2$}           &                                  & +0.43 (significant)             & +0.43                     &                           &                              \\
    \textit{$\Delta_3$}       &                                  & +0.30 (significant)             & +0.66                     &                           &                              \\
\textit{$\Delta_4$}           &                                  &                                 &                           & +0.05                     &                             
\end{tabular}
}
\caption{Comparison of the studies}
\label{synthesis_table}
\end{table*}

Coming back to our initial hypotheses, the $\Delta_2$ results in studies 2 and 3
reinforce both $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The \emph{interaction
hypothesis} $\mathcal{H}_{3}$ was that the activity of modelling one's partner
is not simply the result of an individual cognitive effort but overall the
side-effect of the quality of interactions. The \emph{second-order-modelling
hypothesis} $\mathcal{H}_{2}$ stated that, if $A$ is accurate in modelling $B$,
$A$ should notice when $B$ misunderstands $A$ and should be able to repair. A
good modeler would somehow drive the modelling of his or her partner.  This
\emph{drive} will however occur through verbal cues and repairs and is hence not
easy to differentiate from $\mathcal{H}_{3}$.  The contradictory $\Delta_1$
results in studies 4 (to some extent) and 5 weaken the hypothesis.
